Linear Algebra as a Unifying Theory\\of Wireless Workload Modeling

### 2008-02-02 paper.tex
Work on the paragraph introducint our standard model.

The {packets-bytes-duration} model representation is clearly quite simplistic, not to mention incomplete. It captures nothing about variability in packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; and it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to aggregate behaviors of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: if the end-points, start time and end time of each flow are known, it suffices to know empirical distributions of packet sizes and inter-packet intervals for each flow to reproduce network performance accurately. Accordingly, we represent a flow's behavior by these precise characteristics. We encode a single flow's behavior with six component vectors:
\begin{itemize}
\item $\vec{z} \in \R^{d_z}$ is a histogram of packet sizes;
\item $\vec{v} \in \R^{d_v}$ is a histogram of inter-packet intervals;
\item $\vec{b} \in \R^{d_t}$ indicates when the flow begins;
\item $\vec{e} \in \R^{d_t}$ indicates when the flow ends;
\item $\vec{s}\in\R^n$ indicates the source node of the flow;
\item $\vec{d}\in\R^n$ indicates the destination node of the flow.
\end{itemize}
The behavior vectors for all flows are collected into six corresponding matrices, whose rows are the behavior vectors above:
\begin{itemize}
\item $Z\in\R^{f \times d_z}$ the size matrix;
\item $V\in\R^{f \times d_v}$ the interval matrix;
\item $B\in\R^{f \times d_t}$ the start time matrix;
\item $E\in\R^{f \times d_t}$ the end time matrix;
\item $S\in\R^{f \times n}$ the source matrix;
\item $D\in\R^{f \times n}$ the destination matrix.
\end{itemize}

%We collect all the components of flow behavior into matrix representations of behavior across an entire traffic trace. If there are $f$ flows in the trace, we define the size matrix $Z\in\R^{f \times d_z}$ to be the matrix whose rows are the size vectors $\vec{z}_i$ for $i=1,\dots,f$. In precisely the same manner, we define the interval matrix, $V\in\R^{f \times d_v}$, and the start time matrix, $A\in\R^{f \times d_t}$, and the end time matrix, $E\in\R^{f \times d_t}$.

>>>>

The {packets-bytes-duration} model representation is clearly overly simplistic and very incomplete. (Yet, it is more detailed than most commonly used experimental traffic models!) It captures nothing about variability of packet sizes, nor about the ``burstiness'' of packet transmissions; it leaves start time unspecified; it says nothing about the source or destination of the flow. At the opposite extreme, it is possible to embed the exact start time of each flow, and the exact size and transmission times of its packets in a vector space, etc. The difficulty comes in doing so in a manner that makes the summation of vectors correspond naturally and meaningfully to the aggregate behavior of flows. We take an intermediate approach, suggested by the results of~\cite{Karpinski07:realism}: to reproduce network performance accurately, it suffices to know, for each flow, the correct source node, destination node, start time, end time, and empirical distributions of packet sizes and inter-packet intervals for each flow. Accordingly, we represent a flow's behavior by these precise characteristics. We encode a single flow's behavior with six vectors:
\begin{itemize}
\item $\vec{s}\in\R^n$ indicates the source of the flow;
\item $\vec{d}\in\R^n$ indicates the destination of the flow.;
\item $\vec{b} \in \R^{d_t}$ indicates when the flow begins;
\item $\vec{e} \in \R^{d_t}$ indicates when the flow ends;
\item $\vec{z} \in \R^{d_z}$ is a histogram of packet sizes;
\item $\vec{v} \in \R^{d_v}$ is a histogram of inter-packet intervals.
\end{itemize}
If there are $f$ flows in the entire workload pattern, then its collective behavior can be described with six matrices:
\begin{itemize}
\item $S\in\R^{f \times n}$ the source matrix;
\item $D\in\R^{f \times n}$ the destination matrix;
\item $B\in\R^{f \times d_t}$ the start time matrix;
\item $E\in\R^{f \times d_t}$ the end time matrix;
\item $Z\in\R^{f \times d_z}$ the size matrix;
\item $V\in\R^{f \times d_v}$ the interval matrix.
\end{itemize}
The rows of each matrix are behavior vectors of the $f$ flows in the workload, indexed in some chosen order.
